---
title: "README"
author: "Clara Abdurazak | Getting and Cleaning Data Course Project"
date: "January 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-----

### **README: Analyzing Wearable Computing Data**

This README file describes the files and scripts in the `GettingandCleaningData` repo, how they all work and how they are connected. 

View HTML version: http://htmlpreview.github.io/?https://github.com/clarablara/GettingandCleaningData/blob/master/README.html

-----

### Included Files

This repo includes the following files:

- `\data` - Folder containing the raw (Samsung) data sets. Please see the Code Book for more information on the raw data.

- `CodeBook.Rmd` - R markdown file that describes the variables, data (including units), transformations and calculations performed to clean up and summarize the Samsung wearable computing data 

- `CodeBook.html` - HTML version of the Code Book

- `README.Rmd` - R markdown file that describes the files and scripts in the `GettingandCleaningData` repo, how they all work and how they are connected

- `README.html` - HTML version of the README

- `run_analysis.R` - R script that performs analysis on the raw data set (contained within the `\Data` folder) to clean up and summarize the wearable computing data. (See below for more information on `run_analysis.R`.)

- `tidy_data.txt` - A text file that contains a tidy data set with the average of each variable for each activity and each subject analyzed from the raw data set. (See below for more information on `tidy_data.txt`.)


-----


### Important Assumptions

The `run_analysis.R` script will run correctly as long as the Samsung data is in your working directory within a folder named `\data`.  

The script requires installation the `{dplyr}` R package on your machine. The script will automatically load `{dplyr}` as long as it is installed correctly.

**Please ensure that the following Samsung data files are in your `\data` folder before running the script:**

- `\test` - a folder containing the files: `subject_test.txt`, `X_test.txt`, `Y_test.txt`

- `\train` - a folder containing the files: `subject_train.txt`, `X_train.txt`, `Y_train.txt`

- `features.txt`

*(The other files contained in the Samsung data set are not necessary for this analysis.)*

-----


### Description of `run_analysis.R` Script

The `run_analysis.R` script runs an analysis on the raw (Samsung) data set to clean up and summarize the wearable computing data. 

The script includes the following five (5) functions:

1. `merge_data()` - Merges the training and the test sets to create one data set;  

2. `extract_data()` - Extracts the measurements on the mean and standard deviation for each measurement;  

3. `activity_names()` - Sets descriptive activity names to name the activities in the data set;  

4. `variable_names()` - Labels the data set with descriptive variable names (for features);  

5. `tidy_data()` - Creates tidy data set with the average of each variable for each activity and subject.  
    - NOTE: The `tidy_data()` function makes use of `{dplyr}` package

-----


### Sourcing and Output

Sourcing the `run_analysis.R` script from your working directory will automatically call all the necessary functions to clean, analyze and summarize the raw data. The output will be a tidy data set, `tidy_data.txt`, which will be created in your working directory.

If the script runs correctly, you will see the following messages on your console:

```{r, eval=FALSE}
Running run_analysis.R program ...
... Working on Step 1: Reading and merging Samsung data sets from "data" folder ...
... Working on Step 2: Extracting mean and standard deviation measurements from raw data ...
... Working on Step 3: Labelling data set with descriptive activity names ...
... Working on Step 4: Labelling data set with descriptive variable names ...
Success! The file "tidy_data.txt" has been created in your working directory.
```


The output `tidy_data.txt` file contains a tidy data set with the average of each variable for each activity and each subject analyzed from the raw (Samsung) data set. The dimensions of `tidy_data.txt` are: 180 observations (rows) by 68 variables (columns).

After running the script, you can read the `tidy_data.txt` file into R with the following command:

```{r, eval=FALSE}
read.table("tidy_data.txt", header = TRUE)
```

Please see the Code Book for a more complete description of variables and data in `tidy_data.txt`.

-----
